### 2023-10-24 main.tex
Re-organized LaTeX files
Moved title information to manuscript.tex
Moved figures to after references

\documentclass[11pt,letterpaper]{article}

% PACKAGE CONFIGURATIONS
\usepackage{amsmath}
\usepackage{graphicx}

% Margins and spacing
\usepackage[margin=1in]{geometry}
\usepackage{lscape}
%\usepackage{setspace}  % double spacing
%\doublespacing
\usepackage[skip=10pt]{parskip}  % Remove indents
\setcounter{secnumdepth}{0}  % Remove section numbering

% Figure formatting
\usepackage{longtable}
\usepackage[labelfont=bf, labelsep=period]{caption}
\renewcommand{\figurename}{Fig}

% References
\usepackage[backend=biber, style=nature, doi=false, url=false]{biblatex}
\addbibresource{refs.bib}
\usepackage{hyperref}

% DOCUMENT
\begin{document}

\graphicspath{{figures/}}
\input{manuscript.tex}

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}

\clearpage
\section*{Supporting information}
\input{supporting.tex}

\end{document}